---
title: "Boot Camp Final Project"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# You are a data analyst for the New York City Department of Education. You've been tasked with answering the question:

Q: What can data tell us about the relationship between poverty and test performance in New York public schools?

A: 

Some other questions that your department has:
1. What's the difference in test performance between low, medium and high poverty areas?
2. Has this relationship changed over time?
3. Is this relationship at all moderated by access to free / reduced price lunch?
```{r}
library(dplyr)
library(data.table)
library(ggplot2)
```


## STEP1 Clean the data
# import ur data and explore data
```{r chunk name}

nys_schools <- read.csv("nys_schools.csv")
nys_acs <- read.csv("nys_acs.csv")
summary(nys_schools)
summary(nys_acs)
```


```{r}
#deal with missing values, -99
nys_schools[nys_schools == -99]  <- NA
summary(nys_schools)
```

```{r}
#create a categorical variable groups counties
hist(nys_acs$county_per_poverty)


quantiles <- quantile(nys_acs$county_per_poverty,c(0,1/3,2/3,1))
# Create the Poverty_Group variable
nys_acs$Poverty_Group <- cut(nys_acs$county_per_poverty, 
                             breaks = quantiles, 
                             labels = c("Low", "Medium", "High"))

head(nys_acs)
```

```{r}
#create new variable standardized z-score for ELA for each year
nys_schools <-nys_schools %>%
  group_by(year) %>%
  mutate(
    Math_Z_Score = scale(mean_math_score),
    Ela_Z_Score = scale(mean_ela_score)
  )

```


```{r}
total <- merge(nys_acs,nys_schools,by.nys_acs = c("county_name","year") , by.nys_schools = c("county_name","year"))
#total <- inner_join(nys_acs,nys_schools,by = "year")

summary(total)
#inner join delete, left join , outer join all 
```




## STEP 2 Analyze the data

```{r}
#create a summary tables
percent_in_poverty = data.frame()

```


```{r}
#data visualization
plot(total$Poverty_Group ~ total$ ) %>% group_by(county_name)
```


